Stack Safety as a Security Property

### 2020-07-10 main.tex
Remove pinit

We now present our model of enforcement, which we introduced informally in \Cref{sec:threat}.
A policy is a general model of an enforcement mechanism, consisting of a
set of policy states \(\pol \in \POLS\), a policy step function
\((\mach, \pol) \polstep \pol' : \MACHS \times \POLS
\rightharpoonup \POLS\), and an initializer \(\pinit : \MACHS \rightharpoonup
\POLS\). The initializer establishes the initial policy state for a given
machine state, but is partial to reflect that some policies are only defined
on states that are well-formed in some way.
%
\apt{Is this our current thinking?}
\leo{I thought this was what we settled on in the last two days'
  discussions}\bcp{That was not my impression!  I don't see how the policy
  state can be recovered from just the machine state---e.g., the machine
  state doesn't say where the calls and returns are, does it?!}
%
A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
\Cref{sec:enforcement}.

@@

We say that a system enjoys {\em stack integrity} if for every initial
state $\mach_0$ and its corresponding induced trace $\MPT$ (with
$\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow \MPT$), for every
$n$-transition from $\mpstate{\mach_c}{\pol_c}$ to
$\mpstate{\mach_e}{\pol_e}$ in $\MPT$, if we take the prefix of the
call until its corresponding return $\mpstate{\mach_e}{\pol_e}
\hookrightarrow \MPT_{call} ~ | ~ \ret{\mach_c}$, then $\MPT_{call}$ must
satisfy trace integrity:
\[\propdef{TraceInt} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\]

@@

\definition%{Stack Confidentiality}
%Let $\callmap$ be a call map and $\codemap$ a function map.
For every initial state \(\mach_0\), we take the induced trace
\(\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow \MPT\). Then for
any \(n\)-transition from $\mpstate{\mach_c}{\pol_c}$ to
$\mpstate{\mach_e}{\pol_e}$ in $\MPT$, we take the prefix of the trace
from the entry point until its corresponding return:
%
\[\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c}\]
%
We also take the corresponding trace from any variant \(\mach_e'
\approx_{\mathit{Cof}(\mach_c,n)} \mach_e\):
%
\[\mach_e' \hookrightarrow \machT_{\mathit{call}} | \ret{\mach_c}\]

@@

If trace confidentiality holds for all such traces,
\(\propdef{TraceConf} ~ \ret{\mach_c} ~ \MPT_{\mathit{call}} ~
\machT_{\mathit{call}}\) and
\(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\), then the
system as a whole enjoys {\em stack confidentiality} with respect to
\(\callmap\).

@@

\definition%{Control Separation}
A system enjoys {\em control separation}
% with respect to a function map
%\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\),
if, for any initial state \(\mach_0\) and any adjacent pair of states
\(\mpostate{\mach_1}{\pol_1}{\obs_1}\) and \(\mpostate{\mach_2}{\pol_2}{\obs_2}\)
in the induced trace from \(\mpstate{\mach_0}{\pinit(\mach_0)}\),
if \(\codemap(\mach_1(\PCname)) \not = \codemap(\mach_2(\PCname))\),
then \(\mpstate{\mach_1}{\pol_1}\) is either a call \(\callmap(\mach_1(\PCname))\)
or a return \(\retmap(\mach_1(\PCname))\).

@@

\definition%{Entry Integrity}
A system enjoys {\em entry integrity}
%with respect to a call map
%\(\callmap\) and entry map \(\entmap\)
if, for any initial state \(\mach_0\)
and any n-transition from \((\mach_c,\pol_c)\) to
\((\mach_e, \pol_e)\) in the induced trace from
\(\mpstate{\mach_0}{\pinit(\mach_0)}\), $\mach_e$ is a valid entry:
\[\callmap(\mach_c(\PCname)) = n \Rightarrow \entmap(\mach_e)\]

@@

\definition%{Return Integrity}
A system enjoys {\em return integrity} if all of its returns match prior calls
and return to the appropriate location. For any initial state \(\mach_0\), let
\(\MPT\) be the induced trace of \(\mpstate{\mach_0}{\pinit(\mach_0)}\).
\(\MPT\) may not have an unmatched return -- \(\Unmatched ~ \MPT ~ \MPT_{tail}\)
may not hold for any \(\MPT_{tail}\). For any state pair \((\mach_c, \pol_c)\)
corresponding to a call \(\MPT\), if
\(\mpstatename = \mpostate{\mach_r}{\pol_r}{\obs_r}\)
is the first unmatched return such that
\( \Unmatched ~ \MPT ~ \MPT_{tail}\) and \(\head{\MPT_{tail}} = \mpstatename\),
then $\mach_r$ must correspond to a return state for $\mach_c$:
\(\ret{\mach_c}\ \mach_r\)

@@

\definition
%
A system satisfies {\em lockstep stack safety} if for every initial machine $\mach$,
the trace induced from $(\mach, \pinit(\mach))$ satisfies trace lockstep stack safety with
respect to the empty stack.

@@

      {\em Observable stack confidentiality} is structured like stack
      confidentiality. Let \(\callmap\) be a callmap; for each initial
      state \(\mach_0\) and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
      to \(\mpstate{\mach_e}{\pol_e}\) in the induced trace
      \(\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow \MPT\), take any
      variation \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} |
      \ret{\mach_c}\) and \(\mach_e' \hookrightarrow \machT_{\mathit{call}} |
      \ret{\mach_c}\). We must have \(\propdef{ObsTraceConf} ~
      \ret{\mach_c} ~ \MPT ~ \machT\).

>>>>

We now present our model of enforcement, which we introduced
informally in \Cref{sec:threat}.  A policy is a general model of an
enforcement mechanism, consisting of a set of policy states \(\pol \in
\POLS\), a policy step function \((\mach, \pol) \polstep \pol' :
\MACHS \times \POLS \rightharpoonup \POLS\). We assume that the
compiler provides an initial policy state $\pol_0$ to accompany 
the initial machine state.
%
A concrete policy based on \citet{DBLP:conf/sp/RoesslerD18} will be described in
\Cref{sec:enforcement}.

@@

We say that a system enjoys {\em stack integrity} if for every initial
state $(\mach_0,\pol_0)$ and its corresponding induced trace $\MPT$ (with
$\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT$), for every
$n$-transition from $\mpstate{\mach_c}{\pol_c}$ to
$\mpstate{\mach_e}{\pol_e}$ in $\MPT$, if we take the prefix of the
call until its corresponding return $\mpstate{\mach_e}{\pol_e}
\hookrightarrow \MPT_{call} ~ | ~ \ret{\mach_c}$, then $\MPT_{call}$ must
satisfy trace integrity:
\[\propdef{TraceInt} ~ \mathit{Cof}(\mach_c,n) ~ \MPT_{call}\]

@@

\definition%{Stack Confidentiality}
%Let $\callmap$ be a call map and $\codemap$ a function map.
For every initial state \((\mach_0,\pol_0)\), we take the induced trace
\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT\). Then for
any \(n\)-transition from $\mpstate{\mach_c}{\pol_c}$ to
$\mpstate{\mach_e}{\pol_e}$ in $\MPT$, we take the prefix of the trace
from the entry point until its corresponding return:
%
\[\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} ~ | ~ \ret{\mach_c}\]
%
We also take the corresponding trace from any variant \(\mach_e'
\approx_{\mathit{Cof}(\mach_c,n)} \mach_e\):
%
\[\mach_e' \hookrightarrow \machT_{\mathit{call}} | \ret{\mach_c}\]

@@

If trace confidentiality holds for all such traces,
\(\propdef{TraceConf} ~ \ret{\mach_c} ~ \MPT_{\mathit{call}} ~
\machT_{\mathit{call}}\) and it also holds for the toplevel execution
trace \(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\), then the
system as a whole enjoys {\em stack confidentiality} with respect to
\(\callmap\).

@@

\definition%{Control Separation}
A system enjoys {\em control separation}
% with respect to a function map
%\(\codemap\), a call map \(\callmap\), and a return map \(\retmap\),
if, for any initial state \((\mach_0, \pol_0)\) and any adjacent pair of states
\(\mpostate{\mach_1}{\pol_1}{\obs_1}\) and \(\mpostate{\mach_2}{\pol_2}{\obs_2}\)
in the induced trace from \(\mpstate{\mach_0}{\pol_0}\),
if \(\codemap(\mach_1(\PCname)) \not = \codemap(\mach_2(\PCname))\),
then \(\mpstate{\mach_1}{\pol_1}\) is either a call \(\callmap(\mach_1(\PCname))\)
or a return \(\retmap(\mach_1(\PCname))\).

@@

\definition%{Entry Integrity}
A system enjoys {\em entry integrity}
%with respect to a call map
%\(\callmap\) and entry map \(\entmap\)
if, for any initial state \((\mach_0, \pol_0)\)
and any n-transition from \((\mach_c,\pol_c)\) to
\((\mach_e, \pol_e)\) in the induced trace from
\(\mpstate{\mach_0}{\pol_0}\), $\mach_e$ is a valid entry:
\[\callmap(\mach_c(\PCname)) = n \Rightarrow \entmap(\mach_e)\]

@@

\definition%{Return Integrity}
A system enjoys {\em return integrity} if all of its returns match prior calls
and return to the appropriate location. For any initial state \((\mach_0\, \pol_0)\), let
\(\MPT\) be the induced trace of \(\mpstate{\mach_0}{\pol_0}\).
\(\MPT\) may not have an unmatched return -- \(\Unmatched ~ \MPT ~ \MPT_{tail}\)
may not hold for any \(\MPT_{tail}\). For any state pair \((\mach_c, \pol_c)\)
corresponding to a call \(\MPT\), if
\(\mpstatename = \mpostate{\mach_r}{\pol_r}{\obs_r}\)
is the first unmatched return such that
\( \Unmatched ~ \MPT ~ \MPT_{tail}\) and \(\head{\MPT_{tail}} = \mpstatename\),
then $\mach_r$ must correspond to a return state for $\mach_c$:
\(\ret{\mach_c}\ \mach_r\)

@@

\definition
%
A system satisfies {\em lockstep stack safety} if for every initial machine $(\mach,\pol)$,
the trace induced from $(\mach, \pol)$ satisfies trace lockstep stack safety with
respect to the empty stack.

@@

      {\em Observable stack confidentiality} is structured like stack
      confidentiality. Let \(\callmap\) be a callmap; for each initial
      state \((\mach_0, \pol_0)\) and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
      to \(\mpstate{\mach_e}{\pol_e}\) in the induced trace
      \(\mpstate{\mach_0}{\pol_0} \hookrightarrow \MPT\), take any
      variation \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} |
      \ret{\mach_c}\) and \(\mach_e' \hookrightarrow \machT_{\mathit{call}} |
      \ret{\mach_c}\). We must have \(\propdef{ObsTraceConf} ~
      \ret{\mach_c} ~ \MPT ~ \machT\).
